Concurrent Programming in Harmony

### 2021-11-12 book/paper.tex
First steps towards concurrent model checker

First we can check if the queue implementation
in \autoref{fig:queue} meets the sequential
queue specification in \autoref{fig:queuespec}.
To check if the queue implementation meets the specification,
we need to see if any sequence of queue operations in the
implementation matches a corresponding sequence in the
specification.
We say that the implementation and the specification
have the same \emph{behavior}.
%
\autoref{fig:qtestseq} presents a test program that does
exactly this, for sequences of up to \texttt{NOPS}
queue operations.
It maintains two queues:
\begin{itemize}
\item[] \textit{implq}: the queue of the implementation;
\item[] \textit{specq}: the queue of the specification
\end{itemize}
For each operation, the code first decides whether to
perform a \texttt{get} or \texttt{put} operation.
In the case of a \texttt{put} operation, the code also
decides which value to append to the queue.
All operations are performed on both the queue implementation
and the queue specification.
In the case of \texttt{get}, the results of the operation
on both the implementation and specification are checked against one another.

@@

Given the linearization points, \autoref{fig:qtestconc} shows how
linearizability can be tested.
The test program is similar to the sequential test program
(\autoref{fig:qtestseq}) but starts a thread for each operation.
The operations are executed concurrently on the concurrent queue
implementation of \autoref{fig:queuelin}, but they are executed sequentially
on the sequential queue specification of \autoref{fig:queuespec}.
To that end, the test program maintains a global time variable
\textit{qtime}, and each thread waits until the timestamp assigned to
the last concurrent queue operation matches \textit{qtime} before invoking
the sequential operation in the specification.
Afterward, it atomically increments the shared \textit{qtime} variable.
This results in the operations being executed sequentially against the
sequential specification in the same order of the linearization points
of the concurrent specification.

>>>>

First we can check if the queue implementation
in \autoref{fig:queue} meets the sequential
queue specification in \autoref{fig:queueseq}.
To check if the queue implementation meets the specification,
we need to see if any sequence of queue operations in the
implementation matches a corresponding sequence in the
specification.
We say that the implementation and the specification
have the same \emph{behavior}.
%
\autoref{fig:qtestseq} presents a test program that does
exactly this, for sequences of up to \texttt{NOPS}
queue operations.
It maintains two queues:
\begin{itemize}
\item[] \textit{implq}: the queue of the implementation;
\item[] \textit{specq}: the queue of the specification
\end{itemize}
For each operation, the code first decides whether to
perform a \texttt{get} or \texttt{put} operation.
In the case of a \texttt{put} operation, the code also
decides which value to append to the queue.
All operations are performed on both the queue implementation
and the queue specification.
In the case of \texttt{get}, the results of the operation
on both the implementation and specification are checked against one another.

@@

Given the linearization points, \autoref{fig:qtestconc} shows how
linearizability can be tested.
The test program is similar to the sequential test program
(\autoref{fig:qtestseq}) but starts a thread for each operation.
The operations are executed concurrently on the concurrent queue
implementation of \autoref{fig:queuelin}, but they are executed sequentially
on the sequential queue specification of \autoref{fig:queueseq}.
To that end, the test program maintains a global time variable
\textit{qtime}, and each thread waits until the timestamp assigned to
the last concurrent queue operation matches \textit{qtime} before invoking
the sequential operation in the specification.
Afterward, it atomically increments the shared \textit{qtime} variable.
This results in the operations being executed sequentially against the
sequential specification in the same order of the linearization points
of the concurrent specification.
